Artificial Ecosystem Selection Update

### 2020-10-21 Presentation/presentation.Rmd
update heritability analysis and presentation

# Outline

@@

- Show you what I've been working on
- Hear your questions to see if that guides me down any path
- ???

# The Question

## 
One of the fundamental questions in ecology is, "How does the biodiversity
  that we observe influence ecosystem functions?"

@@

##
For certain functions, this works pretty well. For example, the stability of
  primary productivity in temperate plant communities in the face of drought is
  correlated with species richness of plants.

@@

  1. Without a consistent species concept, it is not obvious at what level of
     biological organization to evaluate this relationship. (the ASV? the
     genus? the family?)
  2. It is not always clear which members of a community are members of a
     particular functional group. (This can be determiend morpholigically for
     plants, but must be determined genetically in natural microbial
     communities.)
  3. Microbial communities are immensely diverse and without solutions to the
     previous two problems, sifting through that diversity is challenging.

## When we apply this BEF approach to microbial communities, there are limited
  gains in terms of our knowlegde of the controls on ecosystem function

@@

- One way to think about microbial communities and ecosystem functions is as assemblages of genes within an ecosystem and a trait at the ecosystem level.
- In fact, this often the only way we can measure microbial communities in natural ecosystems.

@@

fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate),
            se = se(estimate)) %>% 
ggplot(mapping = aes(x = passage, y = mean, ymin = mean - se, ymax = mean + se,
                     color = treat)) + 
  geom_pointrange() +
  scale_color_manual(name = 'Treatment', values = c('gray60', 'darkorange2')) +
  theme_bw()
```

@@

>>>>

@@

- Present my results so far and see if you have any questions that might guide future steps.
- Talk about what "heritability" means in a community context.

# How does biodiversity influence ecosystem functions?

@@

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
knitr::include_graphics('draught-stability.png')
```

##
Reasons this works:

1. There's a clear unit of biological organization - the species. 
2. It's obvious what a "plant" is.
3. Communities are relatively simple - (1-25 species).

@@

  1. Without a consistent species concept, it is not obvious at what level of
     biological organization to evaluate this relationship. (the ASV? the
     genus? the family?)
  2. It is not always clear which members of a community are members of a
     particular functional group. Further, in a metabolic network, it's not
     always obvious which functional group is limiting the end product.
  3. Microbial communities are immensely diverse (100s-1000s of ASVs) and
     without solutions to the
     previous two problems, sifting through that diversity is challenging.

@@

- One way to think about microbial communities and ecosystem functions is as assemblages of genes within an ecosystem and a trait at the ecosystem level.
- In fact, this often the only way we can measure microbial communities in natural ecosystems.
- Considering them this way, they resemble an organism with a phenotype. (In a way that is dangerously Clementsian) 
- Using that metaphor, we can think about ways of identifying genes or organisms that are associated with the presence or rate of particular functions.

## Artificial Ecosystem Selection

A *comparative* approach to this problem would be to go out and sample lets of
ecosystems for their functions and their communities. And do lots of
correlations. That's fine, but it's more of a hypothesis generating tool. A
somewhat more *experimental* approach would be to manipulate the community by
generating lots of different communities under very similar environmental
conditions. And then testing which combinations produce the greatest function.

## Artificial Ecosystem Selection

To do this, I generated 24 ecosystems.

@@

fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate),
            se = se(estimate)) %>% 
ggplot(mapping = aes(x = passage, y = mean, ymin = mean - se, ymax = mean + se,
                     color = treat)) + 
  geom_pointrange() +
  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
                     values = c('gray60', 'darkorange2')) +
  labs(x = "Passage", y = 'Flux (-k)') +
  theme_bw()
```

@@

# How *heritable* is soil methane oxidation rate as an ecosystem trait?

Heritability is "the relative importance of heredity in determining phenotypic
values"

"A character can be 'hereditary' in the sense of being **determined by
the genotype**[1] or in the sense of being **transmitted from parents to offspring**[2]."

[1] Broad-sense heritability $= V_G / V_P$

[2] Narrow-sense heritability $= V_A / V_P$

<small>Falconer and MacKay, 1996</small>

$V_G$ is quite are to estimate, but is of considerable theoretical interest. Estimating $V_G$ would answer the question, "How much of an ecosystem function is determined by the genes in that ecosystem vs. the environmetal conditions in that ecosystem?"

$V_A$ is easier to estimate and is of more practical value in population genetics for its use in breeding programs. It is the ratio of breeding value to phenotype value and determines whether selecting a parent for their traits can produce offspring with those traits.

```{r, fig.width=2, fig.height=3, out.width=3e2, out.height=4.5e2}
h2_per_gen <- read_tsv('../Output/h2_per_gen.tsv')

h2_per_gen_sum  <-
  h2_per_gen %>% 
  summarize(h2_med = median(h2), h2_mad = mad(h2))

ggplot(h2_per_gen, aes(x = passage, y = h2)) +
  geom_point() +
  geom_pointrange(aes(x = 2.5, y = h2_med, 
                                  ymax = h2_med + h2_mad,
                                  ymin = h2_med - h2_mad),
  data = h2_per_gen_sum) +
  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                h^2, ")"))) +
  theme_bw()

Each small point is the h2 for that passage and its offspring. The larger central point is the median h2 (+/- median absolute difference) across all four passages.
